---
title: "Asian American Quality of Life: Analysis"
author: "Miguel Fudolig, Luke Cho, Lawrence Kim, Boya Liu"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
---


```{r, output=F}
library(tidyverse)
library(ggplot2)
library(lavaan)
library(car)
```

# Data set  

This data set is from the 2015 Asian American Quality of Life survey. Participants are from Austin, Texas.

## Input data set


```{r}
qol <- read_csv("AAQoL.csv") |> mutate(across(where(is.character), ~as.factor(.x))) |> 
  mutate(`English Difficulties`=relevel(`English Difficulties`,ref="Not at all"),
         `English Speaking`=relevel(`English Speaking`,ref="Not at all"),
         Ethnicity = relevel(Ethnicity,ref="Chinese"))

qol |> DT::datatable()
```
There are 2,609 responses, some with missing data.


# Summary statistics

### Gender

```{r}
qol |> group_by(`Gender`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```
### Age

```{r}
qol |> filter(!is.na(Age)) |> summarize(age=mean(Age),SD=sd(Age),min=min(Age),max=max(Age))
qol |> filter(Age >= 50) |> summarize (n=n())
qol |> filter(Age >= 65) |> summarize (n=n(), median=quantile(Age,0.5))

```

### Ethnicity

```{r}
qol |> group_by(Ethnicity) |> summarize(n=n()) |> mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Marital Status

```{r}
qol |> group_by(`Marital Status`) |> summarize(n=n()) |> mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Living Alone

```{r}
qol |> mutate(alone=if_else(`No One`==0,"With Others","Alone")) |>
  mutate(alone=factor(alone, levels=c("Alone","With Others"))) |> 
  group_by(alone) |> 
  summarize(n=n()) |> 
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Religion

```{r}
qol |> group_by(Religion) |> summarize(n=n()) |> mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Employment

```{r}
qol |> mutate(`Full Time Employment`= ifelse(`Full Time Employment`==0,"No","Yes")) |> 
                group_by(`Full Time Employment`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### US Born

```{r}
qol |>group_by(`US Born`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Duration

```{r}
qol |> filter(!is.na(`Duration of Residency`)) |> 
  summarize(mean=mean(`Duration of Residency`), SD = sd(`Duration of Residency`))
```

### English Speaking and Difficulty

#### Primary Speakers

1 = Primary Speaker, 0 = non-primary speaker
```{r}
qol |> group_by(`Primary Language`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

#### Profiency
```{r}
qol |> group_by(`English Speaking`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

```{r}
qol |> group_by(`English Difficulties`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Familiarity with America and Ethnic Origin

```{r}
qol |> group_by(`Familiarity with America`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

```{r}
qol |> group_by(`Familiarity with Ethnic Origin`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

Comparing Familiarity with Ethnic Origin and America,
```{r}
qol |> group_by(`Familiarity with Ethnic Origin`,`Familiarity with America`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```
#### Identifying to the ethnic community

```{r}
qol |> group_by(`Identify Ethnically`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```
#### Belonging to the ethnic community

```{r}
qol |> group_by(`Belonging`) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```
#### Perceived Discrimination

```{r}
qol |> group_by(Discrimination) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```

### Income

```{r}
qol <- qol |> mutate(Income_median = case_match(Income,"$0 - $9,999"~"Below",
                                         "$10,000 - $19,999" ~"Below",
                                         "$20,000 - $29,999"~"Below",
                                         "$30,000 - $39,999"~"Below",
                                         "$40,000 - $49,999"~"Below",
                                         "$50,000 - $59,999"~"Below",
                                         "$60,000 - $69,999"~"Above",
                                         "$70,000 and over"~"Above",
                                          .default=Income)) |> 
  mutate(Income_median = factor(Income_median, levels=c("Below","Above")))


qol |> group_by(Income) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```
 We pool the following income levels as "Below": 0-59,999, and "Above Median": >60,000. This is based on Austin's median income in 2015, which was found to be 67,195. (https://www.austinchamber.com/blog/02-21-2017-poverty-income#:~:text=Median%20household%20income%20is%20%2455%2C653,%2467%2C195%2C%20plus%20or%20minus%20%241%2C310.)
 
```{r}
qol |> group_by(Income_median) |> summarize(n=n()) |>
  mutate(pct = n/sum(n)*100) |> 
  arrange(desc(n))
```
 
# Analysis

```{r,output=F}
ps <- function(x,dataset=qol){
  x <- enquo(x)
  dataset |> group_by(!!x) |> 
    summarize(n=n()) |> 
    mutate(pct=n/sum(n)*100)
}
```
## Sources of Information summary statistics;

### Family

```{r}
ps(Family)
```
### Close Friends

```{r}
ps(`Close Friend`)
```

### Acquaintances

```{r}
ps(`Acquaintances`)
```

### Health Professionals

```{r}
ps(`Heal Professionals`)
```

### Mobile Apps

```{r}
ps(`Mobile Apps`)
```

### Social Networks

```{r}
ps(`Social Networks`)

```
### Online Communities

```{r}
ps(`Online Communities`)
```


### Health Websites

```{r}
ps(`Health Website`)
```


### Online sources

```{r}
qol |> mutate(online_source = 1*(`Health Website`=="Yes" |
                                   `Online Communities`=="Yes" | 
                                   `Social Networks`=="Yes" | 
                                   `Email`=="Yes")) |> 
group_by(online_source) |> summarize(n=n()) |> mutate(pct=n/sum(n)*100)
```

## Income

### Source of Information association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Family,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.)) |> 
  filter(Family %in%c("Yes","No")) |>  
  mutate(Family=droplevels(Family)) 
glm(Family~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```


### Source of Information: Close Friends association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Close Friend`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Close Friend`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

### Source of Information: Acquaintances association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Acquaintances,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(Acquaintances~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

### Source of Information: Health professionals association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Heal Professionals`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Heal Professionals`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Source of Information: Mobile apps association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Mobile Apps`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Mobile Apps`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Source of Information: Social Networks association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Social Networks`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Social Networks`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Source of Information: Email association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Email,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(Email~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```


### Source of Information: Online Communities association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Online Communities`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Online Communities`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```


### Source of Information: Health Website association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Health Website`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Health Website`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```





### Other sources of information

```{r}
qol |> select(`Health Info Discription`) |> group_by(`Health Info Discription`) |> summarize(n=n())
```

## DIchotomized Income and Age

### Source of Information association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Family,Income_median,Ethnicity,`English Difficulties`,Age) %>% filter(complete.cases(.)) |> 
  filter(Family %in%c("Yes","No")) |>  
  mutate(Family=droplevels(Family)) 
glm(Family~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)

coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```


### Source of Information: Close Friends association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Close Friend`,Income_median,Ethnicity,`English Difficulties`,Age) %>% filter(complete.cases(.))
glm(`Close Friend`~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

### Source of Information: Acquaintances association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Acquaintances,Income_median,Ethnicity,`English Difficulties`,Age) %>% filter(complete.cases(.))
glm(Acquaintances~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

### Source of Information: Health professionals association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Heal Professionals`,Income_median,Ethnicity,`English Difficulties`,Age) %>% filter(complete.cases(.))
glm(`Heal Professionals`~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

### Source of Information: Mobile apps association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(`Mobile Apps`,Income_median,Ethnicity,`English Difficulties`,Age) %>% filter(complete.cases(.))
glm(`Mobile Apps`~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

### Source of Information: Social Networks association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Age,`Social Networks`,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Social Networks`~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

### Source of Information: Email association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Age,Email,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(Email~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```


### Source of Information: Online Communities association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Age,`Online Communities`,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Online Communities`~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```


### Source of Information: Health Website association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> select(Age,`Health Website`,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Health Website`~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

### Source of Information: Online resources association with age andincome after controlling for ethnicity.

```{r}
qol_1 <- qol |> mutate(online_source = 1*(`Health Website`=="Yes" |
                                   `Online Communities`=="Yes" | 
                                   `Social Networks`=="Yes" | 
                                   `Email`=="Yes")) |> 
  select(Age,online_source,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))

glm(as.factor(online_source)~Age+Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
coef(mod1) |> exp() -> OR
confint(mod1) |> exp() -> CI
CI

data.frame(estimate=OR,LCL = CI[,1], UCL = CI[,2])
```

## income: Only people above age 50

### Source of Information association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> filter(Age >=50) |> 
  select(Family,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.)) |> 
  filter(Family %in%c("Yes","No")) |>  
  mutate(Family=droplevels(Family)) 
glm(Family~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

#### Dichotomized Income

```{r}
qol_1 <- qol |> filter(Age >=50) |> 
  select(Family,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.)) |> 
  filter(Family %in%c("Yes","No")) |>  
  mutate(Family=droplevels(Family)) 
glm(Family~Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

### Source of Information: Close Friends association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |>filter(Age >=50) |>
  select(`Close Friend`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Close Friend`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

#### Dichotomized Income

```{r}
qol_1 <- qol |>filter(Age >=50) |>
  select(`Close Friend`,Income_median,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Close Friend`~Income_median+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

### Source of Information: Acquaintances association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |>filter(Age >=50) |>
  select(Acquaintances,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(Acquaintances~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
```

### Source of Information: Health professionals association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> filter(Age >=50) |>
  select(`Heal Professionals`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Heal Professionals`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Source of Information: Mobile apps association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |>filter(Age >=50) |>
  select(`Mobile Apps`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Mobile Apps`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Source of Information: Social Networks association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |>
  filter(Age >=50) |>
  select(`Social Networks`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Social Networks`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Source of Information: Email association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |> filter(Age >=50) |>
  select(Email,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(Email~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```


### Source of Information: Online Communities association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |>
  filter(Age >=50) |>
  select(`Online Communities`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Online Communities`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```


### Source of Information: Health Website association with income after controlling for ethnicity.

```{r}
qol_1 <- qol |>filter(Age >=50) |>
  select(`Health Website`,Income,Ethnicity,`English Difficulties`) %>% filter(complete.cases(.))
glm(`Health Website`~Income+Ethnicity+`English Difficulties`,data=qol_1,family="binomial") -> mod1
summary(mod1)
car::Anova(mod1)
car::vif(mod1)
```

### Other sources of information

```{r}
qol |> filter(Age >=50) |> select(`Health Info Discription`) |> group_by(`Health Info Discription`) |> summarize(n=n())
```


## Health Care Access and Utilization

We will use a model comparison approach to determine whether accounting for Asian ethnicity will improve model performance. The model of choice will be logistic regression for the following response variables: Physical check up, dental check up, urgent care, and folk medicine.

### Physical check up

```{r}
mod2 <- glm(`Physical Check-up`~Income + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Physical Check-up`~Income + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```
The ethnicity model has a lower BIC and residual deviance, which implies better model performance compared to the model without the ethnicity included.


#### Dichotomized Income

```{r}
mod2 <- glm(`Physical Check-up`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Physical Check-up`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```

We now calculate the odds ratios and 95% confidence intervals.

```{r}
OR <- exp(coef(mod3))
OR_CI <- exp(confint(mod3))

OR_df <- data.frame(estimate = OR, LCL=OR_CI[,1], UCL = OR_CI[,2])
OR_df["Age",] <- OR_df["Age",]^10
OR_df


```



### Dental check up

```{r}
mod2 <- glm(`Dentist Check-up`~Income + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Dentist Check-up`~Income + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```
The ethnicity model has a lower BIC and residual deviance, which implies better model performance compared to the model without the ethnicity included.


#### Dichotomized Income

```{r}
mod2 <- glm(`Dentist Check-up`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Dentist Check-up`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```

We now calculate the odds ratios and 95% confidence intervals.

```{r}
OR <- exp(coef(mod3))
OR_CI <- exp(confint(mod3))

OR_df <- data.frame(estimate = OR, LCL=OR_CI[,1], UCL = OR_CI[,2])
OR_df["Age",] <- OR_df["Age",]^10
OR_df


```

### Urgent Care

```{r}
mod2 <- glm(`Urgentcare`~Income + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Urgentcare`~Income + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```
The ethnicity model has a lower residual deviance but higher BIC, which implies that there is no evidence of better model performance when ethnicity is added to the model.


#### Dichotomized Income

```{r}
mod2 <- glm(`Urgentcare`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Urgentcare`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```

We now calculate the odds ratios and 95% confidence intervals.

```{r}
OR <- exp(coef(mod3))
OR_CI <- exp(confint(mod3))

OR_df <- data.frame(estimate = OR, LCL=OR_CI[,1], UCL = OR_CI[,2])
OR_df["Age",] <- OR_df["Age",]^10
OR_df


```
### Folk medicine

```{r}
mod2 <- glm(Folkmedicine~Income + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(Folkmedicine~Income + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```


#### Dichotomized Income

```{r}
mod2 <- glm(`Folkmedicine`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance`,
            data=qol,
            family="binomial")

mod3 <- glm(`Folkmedicine`~Income_median + Age + Gender + `Health Insurance` + `Dental Insurance` + Ethnicity,
            data=qol,
            family="binomial")
summary(mod2)
car::vif(mod2)
car::Anova(mod2) %>% round(3)

summary(mod3)
car::vif(mod3)
car::Anova(mod3) %>% round(3)

anova(mod2,mod3,test="LRT")
data.frame(BIC_mod2=BIC(mod2), BIC_mod3=BIC(mod3)) |> mutate(Diff=BIC_mod3-BIC_mod2)
```

We now calculate the odds ratios and 95% confidence intervals.

```{r}
OR <- exp(coef(mod3))
OR_CI <- exp(confint(mod3))

OR_df <- data.frame(estimate = OR, LCL=OR_CI[,1], UCL = OR_CI[,2])
OR_df["Age",] <- OR_df["Age",]^10
OR_df


```