---
title: "AAQoL machine learning analysis with unbalanced random forest"
author: "Miguel Fudolig"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 3
---


```{r, output=F}
library(tidyverse)
library(ggplot2)
library(lavaan)
library(car)
library(glmnet)
library(randomForestSRC)
library(caret)
library(ggRandomForests)
```


# Data set  

This data set is from the 2015 Asian American Quality of Life survey. Participants are from Austin, Texas.

## Input data set


```{r}
qol <- read_csv("AAQoL.csv") |> mutate(across(where(is.character), ~as.factor(.x))) |> 
  mutate(`English Difficulties`=relevel(`English Difficulties`,ref="Not at all"),
         `English Speaking`=relevel(`English Speaking`,ref="Not at all"),
         Ethnicity = relevel(Ethnicity,ref="Chinese")) |> 
  mutate(Income_median = case_match(Income,"$0 - $9,999"~"Below",
                                         "$10,000 - $19,999" ~"Below",
                                         "$20,000 - $29,999"~"Below",
                                         "$30,000 - $39,999"~"Below",
                                         "$40,000 - $49,999"~"Below",
                                         "$50,000 - $59,999"~"Below",
                                         "$60,000 - $69,999"~"Above",
                                         "$70,000 and over"~"Above",
                                          .default=Income)) |> 
  mutate(Income_median = factor(Income_median, levels=c("Below","Above")))

qol |> DT::datatable()
```



```{r,output=F}
#| echo: FALSE
ps <- function(x,dataset=qol){
  x <- rlang::enquo(x)
  dataset |> group_by(!!x) |> 
    summarize(n=n()) |> 
    mutate(pct=n/sum(n)*100)
}


```

# Demographics

```{r}
ps(Ethnicity)
ps(Gender)
ps(Religion)
ps(`Full Time Employment`)
ps(Income)
ps(`English Speaking`)
ps(`English Difficulties`)
ps(Discrimination)
```


```{r}
qol |> summarize(age_mean = mean(Age,na.rm=T),
                 age_sd = sd(Age,na.rm=T),
                 age_min = min(Age,na.rm=T),
                 age_max = max(Age,na.rm=T))
```



# Source of Information: Family

```{r}
ps(Family)
```

```{r}
rfdata <- qol |> filter(Family %in% c("No","Yes")) |> 
  mutate(Family=droplevels(Family)) |> 
  select(Family, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Health Insurance`,`Dental Insurance`,`Discrimination`) %>%
  # filter(!is.na(Family)) |> 
  na.omit() |>
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Family ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc",method="brf")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
```

### Cross validation in Random Forests (Run when you have time)

```{r}
# myTrainingControl <- trainControl(method = "repeatedcv",
#                                   number = 10,                   
#                                   repeats = 3,                 
#                                   savePredictions = TRUE,
#                                   classProbs = TRUE,
#                                   verboseIter = TRUE,
#                                   search = "grid")
# 
# 
# set.seed(123)
# 
# model_rf <- train(Family~ .,
#                   data=rfdata,
#                   method = 'rf',
#                   metric = "Accuracy",             
#                   trControl = myTrainingControl,       
#                   importance = TRUE                 
#                   )
# 
# varImp(model_rf)
```


## Training/Test set Variable Importance

### Training Importance

```{r}

set.seed(222)
imbal_index <- caret::createDataPartition(rfdata$Family,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Family~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]
# 
# rfsrc(Family~.,data=train, importance="permute", 
#       perf.type="gmean",
#       splitrule="auc",
#       block.size = 10) ->rfobj
rfobj <- imbalanced(Family ~ .,importance=T,data=train,
perf.type = "gmean",splitrule="auc")

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="family_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance = T,outcome="test")
get.imbalanced.performance(test_rf)

test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="family_test_VIMP.png",width=5,height=5,units="in")
```

# Source of Information: Health Professionals

```{r}
ps(`Heal Professionals`)
```

```{r}
rfdata <- qol |> 
  select(`Heal Professionals`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Health Insurance`,`Dental Insurance`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imbalanced(Heal.Professionals ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")->rfobj

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)

```

## Training/Test set Variable Importance

### Training Importance

```{r}
pos<- rfdata |> filter(Heal.Professionals=="Yes")
neg <- rfdata |> filter(Heal.Professionals==0)

set.seed(222)
imbal_index <- createDataPartition(rfdata$Heal.Professionals,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Heal.Professionals~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Heal.Professionals ~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="healthpro_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance=T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="healthpro_test_VIMP.png",width=5,height=5,units="in")
```

# Health Insurance

```{r}
ps(`Health Insurance`)
```

## Random Forest (```randomForestSRC```)

```{r}
#install.packages("randomForestSRC)

rfdata <- qol |> 
  select(`Health Insurance`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imb <- imbalanced(Health.Insurance ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")
print(imb)

get.imbalanced.performance(imb)

var_importance <- imb$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)
  
# Create ggplot for variable importance
importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_minimal()
  
plot(importance_plot)

```

## Training/Test set Variable Importance

### Training Importance

```{r}
set.seed(222)
# ind_pos <- sample(c(0,1), nrow(pos), replace = T, prob = c(0.7, 0.3))
# ind_neg <- sample(c(0,1), nrow(neg), replace = T, prob = c(0.7, 0.3))
# 
# 
# train <- bind_rows(pos[ind_pos==0,],neg[ind_neg==0,])
# test <- bind_rows(pos[ind_pos==1,],neg[ind_neg==1,])

imbal_index <- createDataPartition(rfdata$Health.Insurance,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Health.Insurance~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Health.Insurance ~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="HIns_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance=T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="HIns_test_VIMP.png",width=5,height=5,units="in")
```

# Dental Insurance

```{r}
ps(`Dental Insurance`)
```

## Random Forest (```randomForestSRC```)

```{r}
#install.packages("randomForestSRC)

rfdata <- qol |> 
  select(`Dental Insurance`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imb <- imbalanced(Dental.Insurance ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")
print(imb)

get.imbalanced.performance(imb)

var_importance <- imb$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)
  
# Create ggplot for variable importance
importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_minimal()
  
plot(importance_plot)

```
## Training/Test set Variable Importance

### Training Importance

```{r}

set.seed(222)
imbal_index <- createDataPartition(rfdata$Dental.Insurance,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Dental.Insurance~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Dental.Insurance ~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="DIns_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance=T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="DIns_test_VIMP.png",width=5,height=5,units="in")
```
# Physical Checkup

```{r}
ps(`Physical Check-up`)
```

## Random Forest (```randomForestSRC```)

```{r}
#install.packages("randomForestSRC)

rfdata <- qol |> 
  select(`Physical Check-up`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Health Insurance`,`Dental Insurance`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imb <- imbalanced(Physical.Check.up ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")
print(imb)
plot(imb,plots.one.page = F)
get.imbalanced.performance(imb)

var_importance <- imb$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)
  
# Create ggplot for variable importance
importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_minimal()
  
plot(importance_plot)

```

## Training/Test set Variable Importance

### Training Importance

```{r}

set.seed(222)
imbal_index <- createDataPartition(rfdata$Physical.Check.up,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Physical.Check.up~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Physical.Check.up~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="PChk_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance=T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="PChk_test_VIMP.png",width=5,height=5,units="in")
```

# Dental Checkup

```{r}
ps(`Dentist Check-up`)
```

## Random Forest (```randomForestSRC```)

```{r}
#install.packages("randomForestSRC)

rfdata <- qol |> 
  select(`Dentist Check-up`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Health Insurance`,`Dental Insurance`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imb <- imbalanced(Dentist.Check.up ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")
print(imb)
plot(imb,plots.one.page = F)
get.imbalanced.performance(imb)

var_importance <- imb$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)
  
# Create ggplot for variable importance
importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_minimal()
  
plot(importance_plot)

```

## Training/Test set Variable Importance

### Training Importance

```{r}
set.seed(222)
imbal_index <- createDataPartition(rfdata$Dentist.Check.up,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Dentist.Check.up~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Dentist.Check.up~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="DChk_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance=T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="DChk_test_VIMP.png",width=5,height=5,units="in")
```

# Urgent Care

```{r}
ps(`Urgentcare`)
```

## Random Forest (```randomForestSRC```)

```{r}
#install.packages("randomForestSRC)

rfdata <- qol |> 
  select(`Urgentcare`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Health Insurance`,`Dental Insurance`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imb <- imbalanced(`Urgentcare` ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")
print(imb)
plot(imb,plots.one.page = F)
get.imbalanced.performance(imb)

var_importance <- imb$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)
  
# Create ggplot for variable importance
importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_minimal()
  
plot(importance_plot)

```

## Training/Test set Variable Importance

### Training Importance

```{r}

set.seed(222)
imbal_index <- createDataPartition(rfdata$Urgentcare,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Urgentcare~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(Urgentcare~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="UC_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance=T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="UC_test_VIMP.png",width=5,height=5,units="in")
```

# Folk Medicine

```{r}
ps(`Folkmedicine`)
```

## Random Forest (```randomForestSRC```)

```{r}
#install.packages("randomForestSRC)

rfdata <- qol |> 
  select(`Folkmedicine`, Ethnicity, Age, Gender,Religion, `Full Time Employment`, Income_median, `English Speaking`, `English Difficulties`,`See Family`:`Community Trust`,`Health Insurance`,`Dental Insurance`,`Discrimination`) %>%
  na.omit() |> 
  rename(Employment=`Full Time Employment`,
         EnglishSpeak=`English Speaking`,
         EnglishDiff=`English Difficulties`) |> 
  as.data.frame() |> 
  rename_with(make.names)

imb <- imbalanced(Folkmedicine ~ .,importance=T,data=rfdata,
                    perf.type = "gmean",splitrule="auc")
print(imb)
plot(imb,plots.one.page = F)
get.imbalanced.performance(imb)

var_importance <- imb$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)
  
# Create ggplot for variable importance
importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_minimal()
  
plot(importance_plot)

```

## Training/Test set Variable Importance

### Training Importance

```{r}
set.seed(222)
imbal_index <- createDataPartition(rfdata$Folkmedicine,p=0.8,list=F,times=1)
imbal_train <- rfdata[imbal_index,]
train <- ROSE::ROSE(Folkmedicine~.,
                          data=imbal_train,
                          seed=3)$data
test<- rfdata[-imbal_index,]

# rfsrc(Family~.,data=rfdata, importance="permute", perf.type="gmean",block.size = 10) ->rfobj
rfobj <- imbalanced(`Folkmedicine` ~ .,importance=T,data=train,
                    perf.type = "gmean",splitrule="auc")
print(rfobj)

print(rfobj)
plot(rfobj,plots.one.page = FALSE)
rfobj$importance

var_importance <- rfobj$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance", x = "Variable", y = "Importance") +
  theme_bw()
  
plot(importance_plot)
ggsave(filename="Folk_train_VIMP.png",width=5,height=5,units="in")
```


### Test Set Importance

```{r}
test_rf <- predict.rfsrc(rfobj,newdata=test,importance = T,outcome="test")
get.imbalanced.performance(test_rf)
test_rf$importance

var_importance <- test_rf$importance[, "all"]
var_importance_df <- data.frame(variable = names(var_importance), importance = var_importance*100)

importance_plot <- ggplot(var_importance_df, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "black") +
  coord_flip() +
  labs(title = "Variable Importance (Test)", x = "Variable", y = "Importance") +
  theme_bw()

importance_plot
ggsave(filename="Folk_test_VIMP.png",width=5,height=5,units="in")
```